---
title: "Assignment 1"
author: "Daniel Loza"
date: "02-04-2026"
format: 
    html:
        theme: cerulean
    pdf:
        theme: cerulean
---

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE

# load libraries
library(tidyverse)
library(ggplot2)
library(readr)
library(here)
```

```{r}
# Import the dataset
df <- read_csv(here("data/33842388_Daniel Loza.csv"))
```

# Introduction
In this report, I will visualise the relationship between marketing spend and the purchase amount, to establish whether a correlation exists between the two variables. As the customers are of different segments, I will also explore whether the correlation varies across different customer segments (to control for segment variables). This analysis will provide insights into the effectiveness of marketing spend in driving purchase amounts and whether this relationship is influenced by customer segmentation.

# Research Question
Does a higher marketing spend correlate with a higher purchase amount? Does this correlation differ across different customer segments?

# Dataset Introduction
The dataset used in this analysis contains seven variables: 'customer_id', 'signup_date', 'marketing_spend', 'visits', 'purchase_amount', 'segment' and 'notes'. These have all been tracked across the duration of a recent marketing capmaign. 

Of particular interest for this investigation are the 'marketing_spend' and 'purchase_amount' variables, as they will allow us to explore the relationship between marketing spend and purchase amount. The 'segment' variable will also be important for controlling for customer segmentation in our analysis. 

The data will be cleaned by filtering out any missing values across the analysed columns, and then subsetting the data to only include the relevant variables for this analysis.

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE

# Subset and clean the data
df <- df %>%
  select(visits, purchase_amount, segment) %>%
  drop_na()
```

## Dataset Description
