---
title: "Assignment 1"
author: "Daniel Loza"
date: "02-04-2026"
format: 
    html:
        theme: cerulean
    pdf:
        theme: cerulean
---

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE

# load libraries
library(tidyverse)
library(ggplot2)
library(readr)
library(here)
```

```{r}
# Import the dataset
df <- read_csv(here("data/33842388_Daniel Loza.csv"))
```

# Introduction
In this report, I will visualise the relationship between number of visits and the purchase amount, to establish whether a correlation exists between the two variables. The dataset contains information on customer visits and their corresponding purchase amounts, which will be analyzed to determine if there is a significant correlation between these two factors.

# Research Question
Does a greater number of visits correlate with a higher purchase amount?

# Dataset Introduction
The dataset used in this analysis contains two key variables: 'number_of_visits' and 'purchase_amount'. The 'number_of_visits' variable represents the total number of visits made by customers, while the 'purchase_amount' variable indicates the total amount spent by customers during those visits.

To ensure a valid data set, the data will be cleaned by removing any rows with missing values in the 'number_of_visits' and 'purchase_amount' columns. This step is crucial to maintain the integrity of the analysis and to ensure that the results are based on complete data.

```{r}
#| echo=FALSE, message=FALSE, warning=FALSE

# Subset and clean the data
df <- df %>%
  select(number_of_visits, purchase_amount) %>%
  drop_na()
```

